% !TEX encoding = UTF-8
% !TEX TS-program = xelatex
% !TEX spellcheck = en_GB
% !TEX engine = xelatex
% !TEX root = ../Herbstrith-H10_over_AI.tex
%
% ███    ██ ██    ██ ███    ███ ██████  ███████ ██████      ████████ ██   ██
% ████   ██ ██    ██ ████  ████ ██   ██ ██      ██   ██        ██    ██   ██
% ██ ██  ██ ██    ██ ██ ████ ██ ██████  █████   ██████         ██    ███████
% ██  ██ ██ ██    ██ ██  ██  ██ ██   ██ ██      ██   ██        ██    ██   ██
% ██   ████  ██████  ██      ██ ██████  ███████ ██   ██        ██    ██   ██ ██

In this section I will closely follow Chapter 1 of the German
textbook~\cite{Neukirch2006}. However, the content is also present in the
English reference~\cite[Chap.~2]{Milne2017}, and sometimes the presentation of
this reference will be recited.

\subsection{Number fields and rings of algebraic integers}
% ███    ██ ██    ██ ███    ███        ███████ ██      ██████  ███████
% ████   ██ ██    ██ ████  ████        ██      ██      ██   ██ ██
% ██ ██  ██ ██    ██ ██ ████ ██        █████   ██      ██   ██ ███████
% ██  ██ ██ ██    ██ ██  ██  ██        ██      ██      ██   ██      ██
% ██   ████  ██████  ██      ██ ██     ██      ███████ ██████  ███████

We start with a series of definitions and remind the reader on some important
results from algebraic number theory and commutative algebra. But at first let
us fix an important notation.

Let \(R\) and \(S\) be commutative rings with unity. Let \(φ: R → S\) be a
ring-homomorphism mapping \(1_R\) to \(1_S\), then \(S\) is called an
\emph{\(R\)-algebra} and we write \(a α\) as a short form for \(φ(a) \cdot α\)
(\(a ∈ R\) and \(α ∈ S\)). We are especially interested in the case
where \(R \subseteq S\) and \(φ\) is chosen to be the embedding of \(R\) into
\(S\). In this situation we denote  by \(R[\seq{α}]\) the smallest ring inside
\(S\) containing \(R\) and all \(\seq{α} ∈ S\). Then \(R[\seq{α}]\) contains all
polynomial expressions in \(\seq{α}\) with coefficients in \(R\), i.e.\ all
elements of the form
\[
  \sum_{(\seq[m]{i}) ∈ ℕ^m} a_{i_1, …, i_m} α^{i_1} … α^{i_m},
\]
where only finitely many \(a_{i_1, …, i_m} ∈ R\) are non-zero.

\begin{defin}
  A finite field-extension \(K\) of the rationals \(ℚ\) is called
  \emph{algebraic number field}. This means that \(K\) is a field and at the
  same time a \(ℚ\)-algebra, that is finite-dimensional viewed as a \(ℚ\)-vector
  space. The \emph{degree} \([K : ℚ]\) is the dimension of \(K\) viewed as a
  \(ℚ\)-vector space.
\end{defin}

For convenience we will always assume that \(K\) is a subset of the complex
pane \(ℂ\).

\begin{exam}
  \begin{exlist}
    \item \(ℚ\) is (up to isomorphism) the only algebraic number field of
    degree \(1\).

    \item \(ℚ[√2] = \set{a + b √2 : a, b ∈ ℚ}\) is an algebraic number field
    of degree \(2\). The inverse of \(a + b √2\), where not both \(a\) and \(b\)
    are \(0\), is given by \(\frac{a - b √2}{a^2 - 2 b^2}\).

    \item \(ℚ[√[3]{2}] = \set{a + b √[3]{2} + c √[3]{4} : a, b ∈ ℚ}\) is an
    algebraic number field of degree \(3\).

    \item \(ℚ\left[\e^{\frac{2 i π}{3}}\right] = \set{a + b \e^{\frac{2 i
    π}{3}}}\) is an algebraic number field of degree \(2\).
  \end{exlist}
\end{exam}

If \(K\) is a number field, then the extension \(K/M\) of number fields is in
fact an \emph{algebraic} one. This means that every \(x ∈ K\) is the root of
some polynomial with coefficients in \(M\). We denote by \(μ_{M, x} ∈ M[X]\) the
monic polynomial with root \(x\) dividing (in \(M[X]\)) all other polynomials
with root \(x\). The polynomial \(μ_{M, x}\) is called \emph{minimal polynomial}
of \(x\) over \(M\). By the minimality condition (w.r.t.\ divisibility in
\(K[X]\)) on \(μ_{M, x}\) this polynomial must be irreducible.

In fact, every element \(x\) in \(K\) is algebraic over the rationals \(ℚ\).
Note that the field of all algebraic elements \(\overline{ℚ}\)---called
\emph{algebraic closure} of \(ℚ\)---is however not a number field, as the
extension \(\overline{ℚ} / ℚ\) is (countably) infinite.

\begin{defin}
  Let \(R ⊂ S\) be commutative rings with unity. Then \(α ∈ S\) is called
  \emph{integral} over \(R\) if it is the root of a monic polynomial with
  coefficients in \(R\), i.e.\ if \(α\) satisfies an equation of the form
  \[
    α^n + a_{n-1}α^{n - 1} + … + a_0 = 0
  \]
  for some \(n ≥ 1\) and some \(\seq[n - 1]{a} ∈ R\). If all elements of \(S\)
  are integral over \(R\) then \(S\) is called \emph{integral} over \(R\). If
  \(R\) is an integral domain then \(R\) is \emph{integrally closed} if for all
  elements \(x ∈ \Quot(R)\) being integral over \(R\) implies \(x ∈ R\).
\end{defin}

Of course, when one wants to use tools from algebra some structure on the
considered sets is needed. Thus, the following theorem, implying that if \(α, β
∈ S\) are algebraic over \(S\), so are their sum and product, is very desirable.

\begin{thm}\label{thm:integral closure}
  Let \(R \subseteq S\) be commutative rings with unity. Then the elements of
  \(S\) that are integral over \(R\) form a subring of \(S\).
\end{thm}

Dedekind gave a proof of this theorem using the following proposition.

\begin{pro}\label{pro:characterization of integral elements}
  Let \(R \subseteq S\) be commutative rings with unity. Then \(α ∈ S\) is
  integral over \(R\) if and only if there exists a finitely generated
  \(R\)-module \(M \subseteq S\) such that \(αM \subseteq M\), in fact \(M =
  R[α]\) can be chosen.
\end{pro}

\begin{proof}[Proof of \cref{thm:integral closure}]
  Let \(α, β ∈ S\) be integral over \(R\) and let \(M\) and \(N\) be finitely
  generated \(R\)-modules contained in \(S\) such that \(αM \subseteq M\) and
  \(βN \subseteq N\) hold. We define the product of the two modules as
  \[
    MN := \set{\sum_{i=1}^k m_i n_i \;\middle\vert\;
     k ∈ ℕ, \seq[k]{m} ∈ M, \seq[k]{n} ∈ N}.
  \]
  Clearly, \(MN\) contains \(0\) as \(M\) (and \(N\)) contains \(0\).
  Furthermore, it is closed under addition, and the inverse of an element in
  \(MN\) can be found by inverting all \(m_i ∈ M\) (or \(n_i ∈ N\)) in the sum.
  Thus, \(MN\) forms a subgroup of \(S\). Note that \(a m_i\) is contained in
  \(M\) for all \(a ∈ R\) and \(m_i ∈ M\) since \(M\) is an \(R\)-module. As a
  consequence,
  \[
    a \sum_{i=1}^k m_i n_i = \sum_{i=1}^k \underbrace{a m_i}_{∈ M} n_i ∈ MN
  \]
  for all \(a ∈ R\) and all \(\seq[k]{m} ∈ M\), \(\seq[k]{n} ∈ N\), and we can
  deduce that \(MN\) is an \(R\)-module.

  Let \(\set{\seq[m]{e}} \subseteq M\) generate \(M\) and \(\set{\seq[n]{f}}
  \subseteq N\) generate \(N\). Then it is easily seen that the finite set
  \[
    \set{e_i f_j : 1 ≤ i ≤ m, 1 ≤ j ≤ n} \subseteq MN
  \]
  generates \(MN\).

  We finish the proof by showing that \(αβ\) and \(α ± β\) satisfy \(αβ\, MN
  \subseteq MN\) and \((α ± β) MN\) respectively. Then the proposition implies
  the claim. But this is easy since
  \[
    αβ \sum_{i=1}^k m_i n_i = \sum_{i=1}^k \underbrace{αm_i}_{∈ M} \,
    \underbrace{βn_i}_{∈ N} ∈ MN
  \]
  holds for all \(\seq[k]{m} ∈ M\) and all \(\seq[k]{n} ∈ N\), and
  \[
    (α ± β) \sum_{i=1}^k m_i n_i =
    \sum_{i=1}^k \underbrace{αm_i}_{∈ M}\, n_i ±
      \sum_{i=1}^k m_i \, \underbrace{βn_i}_{∈ N} ∈ MN
  \]
  holds as well.
\end{proof}

Similarly, one can deduce from \cref{pro:characterization of integral elements}
that being integral is a transitive relation on rings. More formally, we have
the following proposition.

\begin{pro}\label{pro:being integral is transitive}
  Let \(R \subseteq S \subseteq T\) be commutative rings with unity. If \(S\) is
  integral over \(R\) and \(α ∈ T\) is integral over \(S\), then \(α\) is
  integral over \(R\).
\end{pro}

The set \(\overline{R}{}^S\) of all elements of \(S\) that are integral over
\(R\) is called \emph{integral closure} of \(R\) in \(S\). By the theorem above
\(\overline{R}{}^S\) is a subring of \(S\).

We will now return our attention from the general case to our specific situation
and consider the elements of \(ℂ\) that are integral over \(ℤ\). These elements
are called \emph{algebraic integers} and the integral closure of \(ℤ\) in \(ℂ\)
is denoted by \(\algint[]\). Given a number field \(K\), we denote by
\(\algint\) the intersection of \(\algint[]\) with \(K\). In other words,
\(\algint\) is the integral closure of \(ℤ\) in \(K\). To emphasize that we are
considering the ring \(ℤ\) and not any \(\algint\), we call \(ℤ\) the ring of
\emph{rational integers}.

We have that \(\overline{ℤ}{}^ℚ = \algint[] ∩ ℚ = ℤ\). Thus, \(ℤ\) is integrally
closed. This follows from a more general result stating that factorial rings are
integrally closed. By \cref{pro:being integral is transitive} this property of
\(ℤ\) extends to all rings of algebraic integers, we have
\(\overline{\algint}{}^K = \algint\). To make the analogue complete we prove
that \(K\) is the fraction field of \(\algint\) (see~\cref{thm:K is the quotient
field of O K}). However, even more is true, as one can choose the denominator in
the quotient to be a rational integer. More precisely, the following holds.

\begin{pro}
  Let \(K\) be a number field and \(\algint\) its ring of algebraic integers.
  For all \(x ∈ K\) there exists a non-zero rational integer \(n ∈ ℤ \setminus
  \set{0}\) such that their product \(nx ∈ \algint\) is an algebraic integer.
\end{pro}

\begin{thm}\label{thm:K is the quotient field of O K}
  The quotient field of \(\algint\) is (isomorphic to) \(K\) for all number
  fields \(K\).
\end{thm}
\begin{proof}
  By the proposition every \(x ∈ K\) can be written as \(x = αn^{-1}\), where
  \(α ∈ \algint\) is an algebraic and \(n ∈ ℤ \setminus \set{0}\) is a rational
  integer. If \(x = βm^{-1}\) is another representation of this form, then \(αm
  = βn\) must hold in \(\algint\) and thus we can embed \(K\) into the quotient
  field \(Quot(\algint)\) by mapping \(x = αn^{-1}\) to the representative \([α,
  n] ∈ \Quot(\algint)\).

  If on the other hand \([α, β] ∈ \Quot(\algint)\) with \(β ≠ 0\) is given,
  then \(α β^{-1}\) an element of \(K\). Thus, there exist \(γ ∈ \algint\) and
  \(n ∈ ℤ \setminus \set{0}\) such that \(γ n^{-1} = α β^{-1}\)---or put
  differently, such that \([γ, n] ∈ \Quot(\algint)\) is in the same equivalence
  class as \([α, β]\). As a consequence, the embedding defined above is
  surjective.
\end{proof}

We can now deduce that an element \(x ∈ K\) is an algebraic integer if and only
if its minimal polynomial \(μ_{ℚ, x}\) has rational integral coefficients.
Indeed, if \(x\) is a root of the monic polynomial \(p ∈ ℤ[X]\) then \(μ_{ℚ,
x}\) divides \(p\) and thus every root of \(μ_{ℚ, x}\) is an algebraic integer
as well. Now decompose
\[
  μ_{ℚ, x}(X) = \prod_{i=1}^n (X - α_i)
\]
then since \(\algint\) is a ring, the minimal polynomial must have coefficients
in \(\algint ∩ ℚ = ℤ\) and the claim is proven. In fact, we can always find an
algebraic integer \(α ∈ K\) that completely determines the number field \(K\).

\begin{thm}[primitive element theorem]\label{thm:primitive element}
  Let \(L/K\) be an extension of number fields then there exists a
  \emph{primitive element} \(α ∈ \algint[L]\) such that \(L = K[α]\). Moreover,
  if \(μ_{K, α} ∈ K[X]\) is the minimal polynomial of \(α\) over \(K\) then the
  degree of \(μ_{K, α}\) and the degree of the field extension \(L/K\) coincide.
  A \(K\)-basis of \(L\) is given by \(\set{1, α, …, α^{n - 1}}\), where \(n =
  [L : K]\).
\end{thm}

Important tools for studying of number fields and algebraic integers are given
by the norm and trace.

\begin{defin}
  For an extension \(L/K\) of number fields and a fixed element \(x ∈ L\) we
  consider the linear transformation \(λ_x : L → L\) defined by \(λ_x(z) = xz\)
  and define the \emph{trace} of \(x\) as
  \[
    Tr_{L/K} (x) := Tr(λ_x)
  \]
  as well as the \emph{norm} of \(x\) as
  \[
    N_{L/K} (x) := \det(λ_x).
  \]
\end{defin}

 By basic facts from linear algebra, we find that the trace \(Tr_{L/K}: L → K\)
 is in fact a homomorphism between the additive groups of the number fields and
 the norm \(N_{L/K}: L^* → K^*\) is in fact a homomorphism between the groups of
 units.

 From the view of Galois theory one can reinterpret the norm and trace as
 follows.

 \begin{thm}\label{thm:norm and trace}
   Let \(L/K\) be an extension of number fields of degree \(n\). Then there
   exist exactly \(n\) embeddings \(\seq{σ}: L → ℂ\) that fix \(K\) point-wise.
   Furthermore, for all \(x ∈ L\) we have that
   \begin{thmlist}
     \item \(μ_{K, x}(X) = \prod_{i = 1}^n (X - σ_i(x))\),
     \item \(Tr_{L/K}(x) = \sum_{i = 1}^n σ_i(x)\), and
     \item \(N_{L_K}(x) = \prod_{i = 1}^n σ_i(x)\).
   \end{thmlist}
 \end{thm}

We call an extension \(L / K\) of number fields \emph{normal extension} if the
embeddings \(σ_i: L → ℂ\) that fix \(K\) point-wise are in fact automorphisms of
\(L\). We do however have the following equivalent characterizations as well.

\begin{pro}
  For an extension \(L/K\) of number fields the following properties are
  equivalent.
  \begin{thmlist}
    \item \(L/K\) is a normal extension.

    \item If an irreducible polynomial \(p ∈ K[X]\) has one root in \(L\), then
    \(p\) splits over \(L\).

    \item \(L\) is the splitting field of some irreducible polynomial \(p ∈
    K[X]\).
  \end{thmlist}
\end{pro}

From the proposition above it is immediate that every extension \(K / ℚ\) of
degree \(2\) is normal. Moving on to degree \(3\) this changes as for instance
\(ℚ[√[3]{2}] / ℚ\) is not normal. The irreducible polynomial \(X^3 - 2 ∈ ℚ[X]\)
has the root \(√[3]{2}\) in \(ℚ[√[3]{2}]\) but both of the other non-real roots
are not contained in the number field. One can however enlarge \(ℚ[√[3]{2}]\) to
obtain a normal extension. More specifically, if \(L/K\) is an extension of
number fields, then there exists (up to isomorphism) a unique number field \(N
\supseteq L\), such that the extension \(N/K\) is normal. We call \(N\) the
\emph{normal closure} of the extension. In fact, if \(N = K[α]\) then \(N\) can
be chosen as the splitting field of \(μ_{K, α}\). Using the normal closure of
\(L/K\) one can show that norm and trace behave well w.r.t.\ towers of field
extensions.

\begin{cor}
  Let \(K ⊂ M ⊂ L\) be a tower of extensions of number fields. Then we have
  \[
    Tr_{L/K} = Tr_{M/K} • Tr_{L/M} \quad \text{and} \quad
    N_{L/K} = N_{M/K} • N_{L/M}.
  \]
\end{cor}

We fix an extension \(L/K\) of number fields and take another look at
\cref{thm:norm and trace}. Then we find for an algebraic integer \(α ∈
\algint[L]\) that its norm \(N_{L/K}(α)\) and trace \(Tr_{L/K}(α)\) are in fact
products and sums of algebraic integers and thus algebraic integers themselves.
Now since norm and trace map are mappings from \(L\) to \(K\) we can deduce that
both \(N_{L/K}(α)\) and \(Tr_{L/K}(α)\) are contained in \(\algint\). In
particular, we find for all \(α ∈ \algint[L]\) that \(N_{L/ℚ}(α)\) and
\(Tr_{L/ℚ}(α)\) are rational integers.

Furthermore, one finds that an algebraic integer \(α \algint[L]\) is a unit if
and only if its norm \(N_{L/K}(α)\) is a unit in \(\algint\). Indeed, if \(β
N_{L/K}(α) = 1\) holds for some algebraic integer \(β ∈ \algint\), we can
rewrite the norm to find that
\[
  1 = α \underbrace{β \prod_{i = 2}^n σ_i(α)}_{∈ \algint[L]}
\]
holds, where \(\id = \seq{σ}\) denote all the complex embeddings of \(L\) that
fix \(K\) point-wise. In the special case that \(K = ℚ\) we find that \(α ∈
\algint[L]\) is a unit if and only if \(N_{L/ℚ}(α)\) is \(±1\).

As a next step we will further investigate the algebraic structure of
\(\algint\).

\begin{thm}
  Let \(L/K\) be an extension of number fields. Then \(\algint[L]\) is a
  finitely generated free \(\algint\)-module.
\end{thm}

We call a module-basis of \(\algint[L]\) over \(\algint\) an \emph{integral
basis} of \(L\) over \(K\). In particular, we can deduce that \(\algint[L]\) is
a finitely generated free \(ℤ\)-module by setting \(K = ℚ\) in the theorem
above. Every integral basis \(\set{\seq{ζ}} ⊂ \algint[L]\) is in fact a vector
space basis of \(L\) over \(K\) as well, thus its cardinality must coincide with
the degree of the extension. Note however, that not every \(K\)-basis of \(L\)
containing only algebraic integers is an integral basis of \(L\) over \(K\). In
full generality it is hard to find an integral basis, but once the basis is
known the structure of \(\algint[L]\) behaves very nicely with respect to
computability, which is the content of the following corollary.

\begin{cor}\label{cor:O K is computable}
  Let \(K\) be a number field and \(\algint\) its ring of algebraic integers.
  Then \(\algint\) is an efficiently presentable and computably categorical
  \(\lang_{ring}\)-structure. \end{cor} \begin{proof} By the theorem \(\algint\)
  is a finitely generated free \(ℤ\)-module. In fact, it even carries a
  \(ℤ\)-algebra structure. Thus, it is efficiently presentable by \cref{ex:Z is
  computable}. Since \(\algint\) is a ring with unity, it is finitely generated
  and as a consequence of \cref{thm:computable categoricity} \(\algint\) is
  computably categorical.
\end{proof}

\subsection{Ideals of \(\algint\)}
% ██ ██████  ███████  █████  ██      ███████
% ██ ██   ██ ██      ██   ██ ██      ██
% ██ ██   ██ █████   ███████ ██      ███████
% ██ ██   ██ ██      ██   ██ ██           ██
% ██ ██████  ███████ ██   ██ ███████ ███████

We view algebraic integers as generalizations of rational integers. Given a
fixed algebraic integer \(α\) one can show using induction on the absolute value
of its norm \(N_{K/ℚ}(α)\) that \(α\) decomposes into a product of irreducible
elements. However, unlike in the case of rational integers this decomposition is
not unique. Indeed, in \(ℚ[i√{5}]\) one can decompose
\[
  21 = 3 \cdot 7 = (1 + i 2 √{5}) \cdot (1 - i 2 √{5}),
\]
where \(3, 7, 1 + i 2 √{5}\) and \(1 - i2 √{5}\) are irreducible and pair-wise
non-associated algebraic integers.%
\footnote{For full details see the first example in~\cite[Chap.~1,
§~3]{Neukirch2006}.}

It was the idea of German mathematician Ernst Eduard Krummer to generalize the
prime decomposition to ‘ideal numbers’. In his view, there should be ideal
primes \(\mathfrak{p}_1, \mathfrak{p}_2, \mathfrak{p}_3\) and \(\mathfrak{p}_4\)
such that
\[
  3 = \mathfrak{p}_1 \mathfrak{p}_2, \quad
  7 = \mathfrak{p}_3 \mathfrak{p}_4, \quad
  1 + 2 √{-5} = \mathfrak{p}_1 \mathfrak{p}_3, \; \text{and} \;
  1 - 2 √{-5} = \mathfrak{p}_2 \mathfrak{p}_4
\]
then
\[
  21 = (\mathfrak{p}_1 \mathfrak{p}_2)(\mathfrak{p}_3 \mathfrak{p}_4) =
    (\mathfrak{p}_1 \mathfrak{p}_3) (\mathfrak{p}_2 \mathfrak{p}_4)
\]
and the decomposition is again unique.

Since divisibility by a fixed number \(n ∈ ℤ\) gives rise to a congruence
relation \(m_1 \equiv m_2 \mod n\) defined by \(n \mid m_1 - m_2\), it is quite
natural---and was indeed carried out by Richard Dedekind---to view these ‘ideal
numbers’ as congruence relations on \(\algint\). Then the equivalence class
containing \(0\) is an \emph{ideal} in the sense of today's algebra and
‘divisibility’ of ideals \(\mathfrak{a}\) by \(\mathfrak{b}\) can be
replaced by the inclusion of sets \(\mathfrak{a} ⊂ \mathfrak{b}\).

Compare this to the well known case of rational integers. Here every ideal is a
principal ideal. Thus, there exist \(α, β ∈ ℤ\) such that \(\mathfrak{a} = (α)\)
and \(\mathfrak{b} = (β)\) and \(α\) is divisible by \(β\) if and only if \((α)
⊂ (β)\).

As with rational integers, one can addition and multiplication of ideals by
\begin{align*}
  \mathfrak{a} + \mathfrak{b} &:=
    \set{α + β : α ∈ \mathfrak{a}, β ∈ \mathfrak{b}}, \text{ and}\\
  \mathfrak{a} \mathfrak{b} &:=
    \set{\sum_{i=0}^n α_i β_i \mid
      n ∈ ℕ, α_0,\seq{α} ∈ \mathfrak{a}, β_0,\seq{β} ∈ \mathfrak{b}}.
\end{align*}
It is easy to prove that sums and products of ideals are again ideals. In fact,
the set of all ideals of \(\algint\) is a monoid with respect to multiplication,
where the neutral element is given by \(\algint = (1)\). However, unlike in the
case of rational integers we have that
\[
  \mathfrak{a} \mathfrak{b} ⊂ \mathfrak{a, b} ⊂ \mathfrak{a} + \mathfrak{b}
\]
and thus that
\[
  \mathfrak{a, b} \mid \mathfrak{a} \mathfrak{b} \quad \text{and} \quad
  \mathfrak{a} + \mathfrak{b} \mid \mathfrak{a, b}.
\]
Before we can further study divisibility of ideals, we need to investigate the
algebraic properties of \(\algint\).

\begin{defin}
  An integral domain \(D\) is called \emph{Dedekind domain} if \(D\) is
  Noetherian, i.e.\ every ideal of \(D\) is finitely generated, algebraically
  closed, and every non-zero prime ideal \(\mathfrak{p} ⊂ D\) is a maximal
  ideal.
\end{defin}

To study the ideals of algebraic integers the following theorem is essential.

\begin{thm}
  Let \(K\) be a number field. Then its ring of algebraic integers \(\algint\)
  is a Dedekind domain.
\end{thm}

Note that for two ideals \((0) \subsetneq \mathfrak{a, b} \subsetneq \algint\),
the sum \(\mathfrak{a} + \mathfrak{b}\) is the smallest (w.r.t.\ set-inclusion)
ideal containing both \(\mathfrak{a}\) and \(\mathfrak{b}\). Indeed, if
\(\mathfrak{c}\) contains \(\mathfrak{b}\) and \(\mathfrak{c}\) then it contains
all sums of elements in \(\mathfrak{a}\) and \(\mathfrak{b}\). As a consequence,
we call \(\mathfrak{a} + \mathfrak{b}\) the \emph{greatest common divisor} of
\(\mathfrak{a}\) and \(\mathfrak{b}\).

Similarly, the intersection \(\mathfrak{a} ∩ \mathfrak{b}\) is the greatest
ideal of \(\algint\) contained in both \(\mathfrak{a}\) and \(\mathfrak{b}\).
Thus, we call \(\mathfrak{a} ∩ \mathfrak{b}\) the \emph{least common multiple}
of the ideals \(\mathfrak{a}\) and \(\mathfrak{b}\). Before we study the role of
prime ideals with respect to this notion of divisibility, an example is in
order.

\begin{exam}
  Consider the ring of rational integers \(ℤ\) and fix two integers \(n_1, n_2
  ∈ ℤ\). We denote by \(d\) their greatest common divisor and by \(m\) their
  least common multiple. As for their principal ideals the following hold
  \begin{align*}
    (n_1) ∩ (n_2) &= n_1 ℤ ∩ n_2 ℤ =
      \set{n : n_1 \mid n \text{ and } n_2 \mid n} =
      \set{n : m \mid n} = (m),\\
    (n_1)(n_2) &=
      \set{\sum_{i=0}^n α_i β_i :
           n ∈ ℕ, α_0,\seq{α} ∈ n_1 ℤ, β_0,\seq{β} ∈ n_2 ℤ} =\\
      &= \set{\sum_{i=0}^n n_1 k_i n_2 ℓ_i :
              n ∈ ℕ, k_0,\seq{k}, ℓ_0,\seq{ℓ} ∈ ℤ} =
        (n_1 n_2),\\
    \intertext{and using Bézout's identity}
    (n_1) + (n_2) &= \set{α + β : α ∈ (n_1), β ∈ (n_2)} =
       \set{n_1 k + n_2 ℓ : k, ℓ ∈ ℤ} = (d).
  \end{align*}
  Thus, in the case of rational integers greatest common divisor and least
  common multiple have their intended meaning if one replaces integers \(n\)
  with their respective principal ideals \((n)\).
\end{exam}

\begin{thm}
  Let \(\algint\) be the ring of of algebraic integers in some number field
  \(K\) and let \(\mathfrak{a} \subsetneq \algint\) be a non-zero ideal. Then
  there exist up to reordering unique prime ideals \(\seq{\mathfrak{p}} ⊂
  \algint\) such that
  \[
    \mathfrak{a} = \mathfrak{p}_1 … \mathfrak{p}_n.
  \]
\end{thm}

Combining multiple occurrences of the same prime ideal in the decomposition
described in the theorem, one writes
\[
  \mathfrak{a} =
    \prod_{\substack{\mathfrak{p} ⊂ \algint\\\mathfrak{p} \text{ prime ideal}}}
      \mathfrak{p}^{ν_{\mathfrak{p}}},
\]
where \(ν_{\mathfrak{p}} ∈ ℕ\) and all but finitely many exponents are zero.%
\footnote{The constructivist reader will be pleased to hear that \(\algint\)
contains only countably many prime ideals and thus one can fix an order of the
set of prime ideals, such that for all ideals \(\mathfrak{a} ⊂ \algint\) all
non-zero exponents \(ν_{\mathfrak{p}}\) appear in a finite initial segment of
the order.}
Using this product notation of the prime decomposition of ideals, one obtains
for
\[
\mathfrak{a} =
  \prod_{\substack{\mathfrak{p} ⊂ \algint\\\mathfrak{p} \text{ prime ideal}}}
    \mathfrak{p}^{ν_{\mathfrak{p}}}
\quad \text{and} \quad
\mathfrak{b} =
  \prod_{\substack{\mathfrak{p} ⊂ \algint\\\mathfrak{p} \text{ prime ideal}}}
    \mathfrak{p}^{μ_{\mathfrak{p}}},
\]
that their greatest common divisor has the factorization
\[
\mathfrak{a + b} =
  \prod_{\substack{\mathfrak{p} ⊂ \algint\\\mathfrak{p} \text{ prime ideal}}}
    \mathfrak{p}^{\min(ν_{\mathfrak{p}}, μ_{\mathfrak{p}})}.
\]
Thus, if \(\mathfrak{a} + \mathfrak{b} = (1)\), we say \(\mathfrak{a}\) and
\(\mathfrak{b}\) are \emph{relative prime}.

If we notice that the product ideals \(\mathfrak{a = \seq{a}}\), where the
\(\mathfrak{a}_i\) are pair-wise relative prime, is equal to the intersection
\[
  \mathfrak{a} = \bigcap_{i = 1}^n \mathfrak{a}_i,
\]
we have all the tools at hand to restate another important property of the
integers.

\begin{thm}[Chinese remainder theorem]\label{thm:Chinese remainder}
  Let \(\seq{\mathfrak{a}} ⊂ \algint\) ideals, which are pair-wise relative
  prime, and let \(\mathfrak{a} = \cap_{i=1}^n \mathfrak{a}_i\). Then the
  following isomorphism of rings holds
  \[
    \algint / \mathfrak{a} \; \cong \;
    \bigoplus_{i = 1}^n \algint / \mathfrak{a}_i.
  \]
\end{thm}

To conclude this subsection we consider the principal ideals \((p)\) generated
by rational primes \(p ∈ ℤ\). It is easy to deduce that the equality
\[
  (21) = (3)(7) = (1 + i 2 √{5}) (1 - i 2 √{5}),
\]
of principal ideals holds in \(K = ℚ[i √5]\) by plugging-in the definition of
products of ideals. Thus, neither \((3)\) nor \((7)\) can be prime ideals in
\(\algint\). We do however know that no other rational prime can divide these
principal ideals. More formally we have
\begin{pro}
  Let \(\mathfrak{p} ≠ (0)\) be a prime ideal in the ring of algebraic integers
  of some number field \(K\). Then there exists a unique rational prime \(p ∈
  ℤ\) such that \(\mathfrak{p}\) divides the principal ideal \((p) = p
  \algint\).
\end{pro}
% TODO cite this theorem (Thm 120, Baxa)
\todo{cite this theorem (Thm 120, Baxa)}

\subsection{Geometry of numbers}
%  ██████  ███████  ██████  ███    ███ ███████ ████████ ██████  ██    ██
% ██       ██      ██    ██ ████  ████ ██         ██    ██   ██  ██  ██
% ██   ███ █████   ██    ██ ██ ████ ██ █████      ██    ██████    ████
% ██    ██ ██      ██    ██ ██  ██  ██ ██         ██    ██   ██    ██
%  ██████  ███████  ██████  ██      ██ ███████    ██    ██   ██    ██

\subsubsection{Minkowski's theorem on convex bodies}

In this section I closely follow \cite[Chap.~I, §2]{Schmidt1991}.

Let \(V\) be an \(n\)-dimensional vector space over the reals \(ℝ\). If \(\seq[k]{e} ∈
V\) (\(k ≤ n\)) are linearly independent over \(ℝ\), then the free Abelian group
\[
  Λ = ℤ e_1 + … + ℤ e_k
\]
is called a \emph{lattice}. Note that \(ℤ + \sqrt{2} ℤ\) is not a lattice in this
sense, because \(1\) and \(\sqrt{2}\) are linearly dependent over \(ℝ\).

Let \(μ\) be the measure corresponding to the usual Euclidean volume%
\footnote{I.e. \(μ\) is the Lebesgue measure on \(ℝ^n\) and therefore the Haar
measure with respect to the locally compact, Abelian group \(⟨Λ, +⟩\).}
on \(ℝ^n\). Then the \emph{fundamental parallelepiped}
\[
  D = \set{\sum_{i=1}^n α_i e_i : α_i ∈ [0, 1]}
\]
has the volume
\[
  μ(D) = |\det \left( \seq{e} \right)|
\]
for a fixed lattice \(Λ = ℤ e_1 + … + ℤ e_n\) in \(ℝ^n\). One calls such a lattice whose rank coincides with the dimension of the vector space \(ℝ^n\) a \emph{full} lattice.

\begin{thm}[Minkowski's theorem on convex bodies] \label{thm:Minkowski}
  Let \(Λ = ℤ e_1 + … + ℤ e_n\) be a full lattice in the \(n\)-dimensional
  \(ℝ\)-vector space \(V\) and let \(D\) denote its fundamental parallelepiped. If \(T
  \subseteq V\) is compact, convex and symmetric in the origin, i.e.\ if \(α ∈ T\)
  so is \(-α ∈ T\), and
  \[
    μ(T) ≥ 2^n μ(D).
  \]
  Then \(T\) contains a non-zero lattice point \(α ∈ Λ \setminus \set{0}\).
\end{thm}

For a proof of Minkowski's theorem and further details on lattices see
\cite[4.4, p.~28]{Neukirch2006} or \cite[Thm.~4.19]{Milne2017}.

\subsubsection{Kronecker's theorem}

I will closely follow the German reference \cite[Chap.~2]{Hlawka1986}. However,
the subject can as well be found in \cite[see][Chap.~\textsc{XXIII}]{Hardy1975}.


\begin{thm}[Dirichlet's unit theorem]\label{thm:Dirichlet}
    see \cite[Thm.~5.1]{Milne2017}
\end{thm}

\begin{thm}[Local-Global--Principle / Hasse-Minkowski theorem]
    see \cite[§ 66]{Meara2000}
\end{thm}

% TODO Fill out missing facts
\todo{Fill out missing facts}
Missing facts
\begin{enumerate}
  %\item Dedekind domain and factorization into prime ideals
  %\item GCD and LCM
  %\item Principal ideals of different primes are relative prime
  % \item
  %  Chinese remainder theorem \cite[see][Chap.~1,
  % Thm~3.6]{Neukirch2006}
  %\item Primitive element theorem
  \item\label{cor:Lagranges four square theorem}
  Lagrange's four-square theorem. A full proof of the claim using
  \cref{thm:Minkowski} is given in \cite[Remark 4.20]{Milne2017}.
  \item abealian extension
  %\item
  Integral bases and computability
  %\item normal extension/closure
  %\item Galois group acts transitively on \(μ_a\).
  %\item self-conjugate elements
\end{enumerate}
